Score-based tests of measurement invariance: Use in practice

### 2013-09-08 frontiers/frontiers.Rnw
add sim1 plots in the file and edit results section roughly
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@250 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[doc]{apa}
\usepackage{graphicx,epsfig,amsmath,alltt,setspace,bm}
\usepackage[english]{babel}

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}
\caption{Get Ordinal Estimate WDMo Alternative Way}
\label{fig:ord1res}
\setkeys{Gin}{width=\textwidth}
<<ord1res, fig=TRUE, height=7, width=7, echo=FALSE>>=
WDMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="WDMo", fplot=TRUE)
#WDMo 
@
\end{figure}

\begin{figure}
\caption{Get Ordinal Estimate maxLMo Alternative Way}
\label{fig:ord2res}
\setkeys{Gin}{width=\textwidth}
<<ord2res, fig=TRUE, height=7, width=7, echo=FALSE>>=
maxLMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="maxLMo", 
               fnl=opval1, fplot=TRUE)
#maxLMo 
@
\end{figure}

@@

\section{Simulation}
\section{Simulation 1}
In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameters such as factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
simulation was to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

@@

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim="sim1", nobs=480, parms=c("loading","error","var"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=",levels(sim1$nlevels),sep="")
  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("ordmax","ordwmax","catdiff")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("ordmax","ordwmax","catdiff"))
@

@@

<<sim2>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
  levels(sim2$test) <- c("ordmax","ordwmax","catdiff")
  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("ordmax","ordwmax","catdiff"))
@

>>>>

\documentclass[doc]{apa}
\usepackage{graphicx,epsfig,amsmath,alltt,setspace,bm,float}
\usepackage[english]{babel}

@@

%Or alternatively, we could use the sctest.lavaan function, but assign ordinal statistics.
\begin{figure}[H]
\caption{Get Ordinal Estimate WDMo Alternative Way}
\label{fig:ord1res}
\setkeys{Gin}{width=\textwidth}
<<ord1res, fig=TRUE, height=7, width=7, echo=FALSE>>=
WDMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="WDMo", fplot=TRUE)
#WDMo 
@
\end{figure}

\begin{figure}[H]
\caption{Get Ordinal Estimate maxLMo Alternative Way}
\label{fig:ord2res}
\setkeys{Gin}{width=\textwidth}
<<ord2res, fig=TRUE, height=7, width=7, echo=FALSE>>=
maxLMo<-sctest(reduced, as.numeric(yg$agegroup),parm=1:4, stat="maxLMo", 
               fnl=opval1, fplot=TRUE)
#maxLMo 
@
\end{figure}

@@

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, pdf=TRUE}
\section{Simulation}
\section{Simulation 1}
In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameters such as factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
simulation was to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

\subsection{Results}
Full simulation results are presented in Figure~\ref{fig:sim11res} to
Figure~\ref{fig:sim13res}; all three test statistics exhibited 
similar results, with the unordered $\max\mathit{LM}$ 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively.

From these figures, one generally observe that the tests isolate 
the parameter violating measurement invariance.  Additionally, the tests 
have somewhat-higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation.

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified.

@@

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim="sim1", nobs=480, parms=c("loading","error","var"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=",levels(sim1$nlevels),sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=",levels(sim1$nobs),sep="")
  levels(sim1$test) <- c("ordmax","ordwmax","catdiff")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("ordmax","ordwmax","catdiff"))
  levels(sim1$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors")

@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Loading1.}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("loading"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\end{figure}

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Covariance}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("var"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\end{figure}

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the ordered double-max test the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is Error1}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7>>=
sim1.tmp <- sim1[sim1$test %in% c("ordmax","ordwmax","catdiff")&sim1$parms %in% c("error"),]
sim1.tmp$test <- factor(sim1.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim1.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\end{figure}

@@

%<<sim2>>=
%## seed for replication
%RNGkind(kind = "default", normal.kind = "default")
%set.seed(1163)

%if(file.exists("sim2.rda")) {
%  load("sim2.rda")
%} else {
%  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))
%  sim2$nlevels <- factor(sim2$nlevels)
%  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
%  sim2$nobs <- factor(sim2$nobs)
%  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
%  levels(sim2$test) <- c("ordmax","ordwmax","catdiff")
%  save(sim2, file="sim2.rda")
%}
%sim2$test <- factor(as.character(sim2$test),
%  levels = c("ordmax","ordwmax","catdiff"),
%  labels = c("ordmax","ordwmax","catdiff"))
%@
